---
title: Proportion of (temperate) reef fish at depth by ontogenetic stage at three different depths ~(surface, mixed layer, below mixed layer) in the top 100m
author: "Steve Hawes"
date: "1 February 2015"
output: html_document
---

## Overview 

Proportions of stages of ontogeny of reef associated fish caught at 3 depths on the RV Investigator 2015

## Importing data and libraries

Libraries used in this code
```{r,warning:FALSE,results:hold}
library("ggplot2") # Graphics 
library("dplyr") # Arranging data
library("reshape2") # Reshaping data
library("RVAideMemoire") # Fisher multiple comparisons
library("ca")
library("vcd")
library("NCStats")
```

My data imported from my excel spreadsheet. I have printed out the first 5 rows so you get the idea. I also grouped the data by the factors of Depth and then Stage.
```{r,results='hold'}
ontogenydata <- read.csv('reef_fish.csv',skip = 2,nrows = 172) ## 172 is the first 16 sites
str(ontogenydata)
names(ontogenydata) <- tolower(names(ontogenydata))
head(ontogenydata)
```

## Functions

1. getFishData()
```{r}
getFishData <- function(df,pre.name,fle.name,post.name){
  fish <- subset(df,select = c("site","depth", pre.name, fle.name, post.name))
  names(fish) <- c("site","depth","preflexion","flexion","postflexion")
  fishmelt <- melt(fish, id.vars = c("site","depth"),
                          measure.vars = c("preflexion","flexion","postflexion"),
                          variable.name = "stage")
  fishmelt <- group_by(fishmelt, depth, stage)
  fishmelt
}
```

2. getProportions() produces the proportions of each stage at each depth of the given data frame. It uses the mutate function from reshape2 to add the proprtions as an extra column
```{r}
getProportions <- function(data){
  data$sum[data$stage == 'preflexion'] = sum(data$count[data$stage == 'preflexion'])
  data$sum[data$stage == 'flexion'] = sum(data$count[data$stage == 'flexion'])
  data$sum[data$stage == 'postflexion'] = sum(data$count[data$stage == 'postflexion'])
  data = mutate(data, prop = count / sum)
  return(data)
}
```

2. graphBuilder() uses ggplot2 to produce a line graph to visualise the differences in proportions. I might change the layout of the graph, its not entirely intuitive at the moment.
```{r,tidy=TRUE}
graphBuilder <- function(data,title){
  
  # Order the factors to print nicely on x-axis
  data$depth <- factor(data$depth,levels=c("0","25","75"))
  
  graph <- ggplot(data, 
                  aes(x = depth, y = prop, fill = stage, group=stage, shape=stage, colour=stage, ymax=1.0, ymin=0)) +
    geom_line(size=1.5) +
    geom_point(size=4) +
    ggtitle(title) +
    scale_y_continuous(breaks = seq(0, 1.0, 0.1), 
                       limits = c(0, 1.0), 
                       expand = c(0, 0)) +
    scale_x_discrete(labels=c("0-3","3-50","50-100")) +
    labs(x = "Depth (m)", y = "Proportion") +
    scale_colour_brewer(palette="Dark2") +
    theme(text = element_text(size=20))
  return(graph)
}
```

3. printLineGraphForFish() is just a placer function to cut down on repeated code
```{r}
printLineGraphForFish <- function(countdata,title) {
  prop <- getProportions(countdata)
  graphBuilder(prop,title)
}
```

4. printMosiacGraphForFish creates mosaic images for presenting Chi-Square tests
```{r}
printMosaicGraphForFish <- function(countdata) {
  matrix <- acast(countdata,depth ~ stage,value.var="count")
  print(matrix)
  plot(ca(matrix))
  mosaic(matrix,shade=T,legend=T)
  assoc(matrix,shade=T,legend=T)
}
```

5. performFishersTest() performs the Fisher's exact test to see if the proportions are significantly different by depth. It uses the reshape2 package's acast() function to convert the data into a matrix format, with rows of Depth and columns of Stage, populated by the count variable. If the two sided fishers exact test is significant, a post-hoc of multiple comparisons using 2x2 fisher's exact tests are used - corrected using a bonferroni approach
```{r,tidy=TRUE}
performFishersTest <- function(countdata) {
  matrix <- acast(countdata,depth ~ stage,value.var="count")
  ft <- fisher.test(matrix,alternative="two.sided")
  ##print(ft)
  if(ft$p.value < 0.05)  {
    mult <- fisher.multcomp(matrix,"bonferroni")
    print(mult)
  }
  return(ft)
}
```

6. performChiSqTest
```{r,tidy=TRUE}
performChiSqTest <- function(countdata) {
  matrix <- acast(countdata,depth ~ stage,value.var="count")
  ft <- chisq.test(matrix)
  ##print(ft)
  if(ft$p.value < 0.05)  {
    mult <- chisqPostHoc(ft,control = "bonferroni")
    print(mult)
  }
  return(ft)
}
```

##Preliminary Results

###Wrasse
It uses the dplyr package to sum all the count data - the totals are needed to calculate the proportions. 
```{r}
labrids <- getFishData(ontogenydata,"pre_labridae","fle_labridae","pos_labridae")
labrids.sum <- summarise(labrids, count = sum(value))
performChiSqTest(labrids.sum)
printLineGraphForFish(labrids.sum,"Labrid")
printMosaicGraphForFish(labrids.sum)
```

###Parrotfish
```{r}
scarids <- getFishData(ontogenydata,"pre_scaridae","fle_scaridae","pos_scaridae")
scarids.sum <- summarise(scarids, count = sum(value))
performFishersTest(scarids.sum)
printLineGraphForFish(scarids.sum,"Scarid")
printMosaicGraphForFish(scarids.sum)
```

###Flounder
```{r}
bothids <- getFishData(ontogenydata,"pre_bothidae","fle_bothidae","pos_bothidae")
bothids.sum <- summarise(bothids, count = sum(value))
performFishersTest(bothids.sum)
printLineGraphForFish(bothids.sum,"Bothid")
printMosaicGraphForFish(bothids.sum)
```

###Scorpionfish
```{r}
scorpids <- getFishData(ontogenydata,"pre_scorpaenidae","fle_scorpaenidae","pos_scorpaenidae")
scorpids.sum <- summarise(scorpids, count = sum(value))
performFishersTest(scorpids.sum)
printLineGraphForFish(scorpids.sum,"Scorpid")
printMosaicGraphForFish(scorpids.sum)
```

###Sea basses & Groupers
```{r}
serranids <- getFishData(ontogenydata,"pre_serranidae","fle_serranidae","pos_serranidae")
serranids.sum <- summarise(serranids, count = sum(value))
performFishersTest(serranids.sum)
printLineGraphForFish(serranids.sum,"Serranid")
printMosaicGraphForFish(serranids.sum)
```

###Lizardfish
```{r}
synodontids <- getFishData(ontogenydata,"pre_synodontidae","fle_synodontidae","pos_synodontidae")
synodontids.sum <- summarise(synodontids, count = sum(value))
performFishersTest(synodontids.sum)
printLineGraphForFish(synodontids.sum,"Synodontid")
printMosaicGraphForFish(synodontids.sum)
```

###Goatfish
```{r}
mullids <- getFishData(ontogenydata,"pre_mullidae","fle_mullidae","pos_mullidae")
mullids.sum <- summarise(mullids, count = sum(value))
performFishersTest(mullids.sum)
printLineGraphForFish(mullids.sum,"Mullid")
printMosaicGraphForFish(mullids.sum)
```

###Damselfish
```{r}
pomacentrids <- getFishData(ontogenydata,"pre_pomacentridae","fle_pomacentridae","pos_pomacentridae")
pomacentrids.sum <- summarise(pomacentrids, count = sum(value))
performFishersTest(pomacentrids.sum)
printLineGraphForFish(pomacentrids.sum,"Pomacentrid")
printMosaicGraphForFish(pomacentrids.sum)
```

